---
title: "HW1_wgeither"
author: "Warren Geither"
date: "8/29/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
Primers Done

## Problem 2
Saved to Github

a) I got my undergrad in mathematics with a minor in stats, so I had a fair amount of exopsure and use in R. For the past 2 years I worked as a Data Analyst at an Ad-tech company where I primarily used Python, SQL, and C#. That being said, in this class I would love to:
 - Refamiliarize myself with linear regression & ANOVA in R
 - Learn about parallel computing and how to connect to the university's supercomputer
 - Learn about Monte Carlo procedures and Power as I'm not sure if this was covered in my undergrad curriculum
 
 b)
 
Binomial Distribution:
 
$$(1)\text{ }P(X=x|n,p) = {{n}\choose{x}} \cdot p^x \cdot (1-p)^{n-x}; x=0,1,2,...,n;\text{ } 0\leq p\leq1$$

Discrete Uniform:

$$(2)\text{ }P(X=x|N) = \frac{1}{N}; x=0,1,2,...,N;\text{ } N\in \mathbb{N}$$

Exponential Distribution:

$$(3)\text{ }f(x|\beta) = \frac{1}{\beta}e^{-x/\beta},\text{ }0\leq x \leq \infty,\text{ } \beta > 0$$

## Problem 3

Objective: plot a scatter plot and histogram of any random R dataset

1.) Look at all available R datasets (I chose Orange)

2.) Read about the dataset and print it out to see what it looks like

3.) Install ggplot2 and initalize library

4.) Plot scatterplot & histogram with ggplot package - use google if you have issues

## Problem 4
```{r}
# look at available R datasets
# library(help="datasets")

# assign stock data set Orange to data variable
data = Orange

# check out the data to see what we can plot
print(data)

# Read about the dataset
# ?Orange
```
```{r}
# install and load package ggplot2 if you did not already
# install.packages("ggplot2")

# initialize library
library("ggplot2")

# use ggplot package to create a scatterplot of age vs. circumference
ggplot(data) +
        geom_point(mapping = aes(x=age, y=circumference), color="blue")
```

```{r}
# use ggplot package to create a histogram of ages of trees (days since 1968/12/31)
ggplot(data) +
        geom_histogram(mapping = aes(x=age), binwidth=50)
```


