---
title: "HW1_wgeither"
author: "Warren Geither"
date: "8/29/2020"
output: pdf_document
header-includes:
 - \usepackage{enumitem}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
Primers Done

## Problem 2 & 3
Saved to Github

a) I got my undergrad in mathematics with a minor in stats, so I had a fair amount of exopsure and use in R. For the past 2 years I worked as a Data Analyst at an Ad-tech company where I primarily used Python, SQL, and C#. That being said, in this class I would love to:
 - Refamiliarize myself with linear regression & ANOVA in R
 - Learn about parallel computing and how to connect to the university's supercomputer
 - Learn about Monte Carlo procedures and Power as I'm not sure if this was covered in my undergrad curriculum
 
 b)
\begin{align}
\text{Binomial: }P(X=x|n,p) &= {{n}\choose{x}} \cdot p^x \cdot (1-p)^{n-x};    x=0,1,2,...,n;\text{ } 0\leq p\leq1 \\
\text{Discrete Uniform: }P(X=x|N) &= \frac{1}{N}; x=0,1,2,...,N;\text{ } N\in \mathbb{N} \\
\text{Exponential: }f(x|\beta) &= \frac{1}{\beta}e^{-x/\beta},\text{ }0\leq x \leq \infty,\text{ } \beta > 0
\end{align}

\textbf{Performing Reproducible Research}
\begin{enumerate}
\item For Every Result, Keep Track of How It Was Produced
\item Avoid Manual Data Manipulation Steps
\item Archive the Exact Versions of All External Programs Used
\item Version Control All Custom Scripts
\item Record All Intermediate Results, When Possible in Standardized Formats
\item For Analyses That Include Randomness, Note Underlying Random Seeds
\item Always Store Raw Data behind Plots
\item Generate Hierarchical Analysis Output, Allowing Layers of Increasing Detail to Be Inspected
\item Connect Textual Statements to Underlying Results
\item Provide Public Access to Scripts, Runs, and Results
\end{enumerate}

## Problem 4
```{r}
# look at available R datasets
# library(help="datasets")

# assign stock data set Orange to data variable
data = Orange

# check out the data to see what we can plot
knitr::kable(summary(data))

# Read about the dataset
# ?Orange
```
```{r}
# install and load package ggplot2 if you did not already
# install.packages("ggplot2")

# initialize library
library("ggplot2")

# use ggplot package to create a scatterplot of age vs. circumference
ggplot(data) +
        geom_point(mapping = aes(x=age, y=circumference), color="blue")
```

```{r}
# use ggplot package to create a histogram of ages of trees (days since 1968/12/31)
ggplot(data) +
        geom_histogram(mapping = aes(x=age), binwidth=50)
```


