---
title: "HW3_wgeither"
author: "Warren Geither"
date: "9/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(curl)
library(ggplot2)
library(microbenchmark)
library(tidyverse)
```

## Problem 3

I already practice alot of the style tips presented. However I can start
to always use double quotes to denote character vectors. 
Since I always used single quotes in python. 


## Problem 5

a.) I can conclude that they have approximately 
    the same mean, variance, and coorelation
```{r}
# load in file from working directory
hw3_data_df <- readRDS("HW3_data (2).rds")

# create function
summarize_dataframe <- function(x){
  
  # group up by the first column and calculate means of col 2 & 3
  mean_summary <- x %>%
                    group_by(x[,1]) %>%
                    summarize(mean_1 = mean(x[,2]),mean_2 = mean(x[,3]))
  
  # same group by first except calc standard deviation for 2&3                      
  stdev_summary <- x %>%
                    group_by(x[,1]) %>%
                    summarize(std_1 = sd(x[,2]),std_2 = sd(x[,3]))
  
  # same except calculating covariance                    
  cor_summary <- x %>%
                  group_by(x[,1]) %>%
                  summarize(cor=cor(x[,2],x[,3]))
  
  # binding all summaries together be suring not to repeat our first col                  
  summary <- cbind(mean_summary
                   ,stdev_summary[,2:3]
                   ,cor_summary[2])
  
  # return final summary
  return(summary)
}

# store result in dataframe
summary_df <- summarize_dataframe(hw3_data_df)

# create knitr table, conclude that they are all pretty much the same
knitr::kable(summary_df)
```

b.) I can conclude that I'm not insane, the data look very similar in terms 
of those values.
```{r boxplots}
# set window
par(mfrow=c(1,2))

# device 1 boxplot
boxplot(dev1~Observer
        , data=hw3_data_df
        , main="Dev1 By Observer"
        , xlab="Observer"
        , ylab="Dev1_values")

# device 2 boxplot
boxplot(dev2~Observer
        , data=hw3_data_df
        , main="Dev2 By Observer"
        , xlab="Observer",
        ylab="Dev2_values")
```

c.) Looking back and forth at both the violin plots, there seems to definitely
   be a difference density at certain points
```{r violin_plot}
# change column to factor column
modified_hw3_data_df <- hw3_data_df
modified_hw3_data_df$Observer <- as.factor(modified_hw3_data_df$Observer)

# graph violin plots
plot_1 <- ggplot(modified_hw3_data_df, aes(x=Observer, y=dev1, color=Observer)) + 
            geom_violin() +
            scale_x_discrete(limits=c(1:12)) +
            ggtitle("Observer By Dev1") +
            labs(y="Dev1", x="Observer")

plot_2 <- ggplot(modified_hw3_data_df, aes(x=Observer, y=dev2, color=Observer)) + 
            geom_violin() +
            scale_x_discrete(limits=c(1:12)) +
            ggtitle("Observer By Dev2") +
            labs(y="Dev2", x="Observer")

# print plot 1
plot_1

# print plot 2
plot_2
```
d.) LOL! Always make sure to look at the data from all angles.
```{r scatterplots}
ggplot(hw3_data_df, aes(x=dev1,y=dev2)) + geom_point() + facet_wrap(Observer~.)
```

## Problem 6 

```{r riemann_approx}
# initialize value for function/loop
sum_val2 = 0

# create function to approximate integral
integral_approx_function <- function(x) {
  
    # loop to get the left rienamm heights and sum them
    for (i in 0:x-1) {
      # input i into function
      sum_val <- exp(-((1/i)^2)/2)
      
      # sum the i's together
      sum_val2 <- sum_val + sum_val2
      }
  
    # muliple by the width of the boxes so we have the total area
    value <- (1/x) * sum_val2
    return(value)
}

# solution is actually sqrt(2pi), just using one though.
numerical_approx <- 1

# initialize other variables
rieman_sum <- 0
delta <- numerical_approx - rieman_sum
j <- 0

# loop through possible values
while (delta > 0.000001){
  # plug j into function
  rieman_sum <- integral_approx_function(j)
  
  # print our values
  print(paste0("Slices being used: ", j))
  print(paste0("Sum calculated: ", rieman_sum))
  
  # increment j
  j <- j + 100000
  
  # find the delta
  delta <- abs(numerical_approx - rieman_sum)
}

```
## Problem 7

```{r newtons_method}
# create function for newtons method
newtons_method_func <- function(x){
  
  # input the equation
  eq <- (3^x) - sin(x) + cos(5*x)
  
  # store the derivative
  derivative_eq <- -5*sin(5*x)-cos(x)+log(3)*3^x
  
  # calculate the next term
  x_n1 <- x - (eq/derivative_eq)
  
  return(x_n1)
}

# initialize guess value
x <- -1

# while difference from root is greater than 0.1 keep going
while (abs(x - 0) > 0.1) {
  
  # find the next value given your guess
  x_n1 <- newtons_method_func(x)
  
  # store that value and continue
  x <- x_n1
  
  # print x
  print(x)
  
}
```

## Problem 8

The matrix operations are much quicker.
```{r calc_SST}
# simulated data
X <- cbind(rep(1,100),rep.int(1:10,time=10))

# create vectore beta
beta <- c(4,5)

# create random vector y
y <- X%*%beta + rnorm(100)

# create mean vector
y_bar <- mean(y)


# initialize values for loop
sum2 <- 0
sum <- 0

# initialize mean matrix
y_bar_vector <- rep(y_bar,100)

# compare finding SST, not sure if theres a better way to wrap the for loop
comparision_table <- microbenchmark(loop = for (i in 1:100) {
  sum <- (y[i] - y_bar)^2
  sum2 <- sum2 + sum
}
, matrix_op = colSums((y- y_bar_vector)^2))

# plot chart
autoplot(comparision_table)

# print table
knitr::kable(head(comparision_table))
```